IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-18 paper/appendix.tex
add some qed's.

\section{IFC Language with a Single Heap}
\label{sec:single-heap}

@@

\section{Extensions}
\label{sec:appendix-extensions}

@@

\section{Limitations and Future Directions}

In this section we state some limitations of the current system, and
outline potential future directions to alleviate these limitations.

\subsection{Type Safety}
\label{sec:extensions:types}

One important consideration for any new language feature is whether
or not it preserves type safety.  If implementations are allowed to
have undefined behavior when the system gets stuck, concerns of type
safety can be directly applicable to security.

In our presentation, we have demanded that implementations \emph{not}
have undefined behavior when getting stuck: in such situations, the
\textsc{I-noStep} rule applies, where the stuck thread should be dropped
from execution.  Our combined language will be not get stuck, even if
the target language could get stuck.  This sort of guarantee can be
achieved at a coarse granularity for languages like C by using hard
isolation.  However, for many other languages, this is too stringent a
requirement.

A possible way to relax this requirement is to drop the \textsc{I-noStep}
rule, and instead demand type-safety of the combined language.  This
depends on the type system of the target language, so it is difficult to
make any general statements about how one would go about doing this.  However,
there are few general remarks to be made:

\begin{itemize}
    \item In Matthews and Findler's original
        paper~\cite{Matthews:2007:OSM:1190216.1190220}, an emphasis was
        on using type boundaries to mediate between the type systems of
        the two languages, i.e. handling conversions of values from
        one language to the other.  These techniques are directly applicable here.

    \item In the case of mini-ES, type safety proceeds by a preservation theorem
        that refers to well-typed stores.  In our setting, there are now multiple
        stores, and expressions may move from being interpreted with one store
        to another (e.g. when an address is sent from one thread to
        another); a clear violation of type safety.  In this case, operations
        such as |send| have their typing rules require
        the expressions being sent be typeable in an arbitrary well-typed store,
        statically disallowing the sending of addresses, or any types which
        may contain addresses.  In practice, this may be further restricted to
        types which can be easily marshalled, e.g. strings.

    \item Similarly, the choice of $\kappa$ (from Section~\ref{sec:retrofit}) directly
        influences the difficulty of the type-safety proof for
        |sandbox| when the target language has a store.  When
        $\kappa$ is identity, most expressions can be allowed, since the store
        remains the same; if $\kappa$ drops the store, however, only closed expressions
        can be sandboxed.
\end{itemize}

\subsection{Scheduling Policies}
Our specification language is parametrized by a scheduling policy |alpha|, which
maps a task list to a potentially different task list, i.e.,
$\alpha_{{\tiny\mathrm{\text{step}}}}(\ifc{\Varid{t}},\ldots)$. 
%
As previously discussed, the precise definition of |alpha|
dictates whether the language is TSNI, TINI, or neither.
%
We described two popular policies |roundrobinf| and
|seqf| for which we showed TSNI and TINI, respectively.
%
However, our definitions allow for other scheduling policies based on the
current task list---in fact, our formalism can be easily 
extended to consider history~\cite{BartheRRS07}.
%
As an example of generality, our approach allows to implement the scheduler of~\cite{Kashyap:2011}
which always schedules less sensitive threads first; this can be
employed to address external timing covert channels where an attacker
can measure the delay between output events.\footnote{
  When considering a label lattice that is not totally ordered, e.g.,
  DLM~\cite{myers:dlm} or DCLabels~\cite{dclabels}, this scheduling policy is
  considerably more complex.
}
%
More precisely, our calculus generalizes to a class of deterministic
schedulers that are: 1) \emph{integrity-preserving}, i.e., they do not
modify (or create) tasks; 2) \emph{non-interferent}, i.e., 
%greatest lower bound of all the labels on the task labels,\footnote{
%  The label on a task's current label is the current label of the
%  parent at the time of creation, i.e., the initial current label of
%  the task.
%} 
scheduling decisions on public tasks do not depend on a sensitive task's
state~\cite{russo2006securing}; and, 3) make \emph{low-progress}, i.e., if there
is a task in the task list whose label is below (or incomparable to) the label
of another task, this task (or another one at the same level) must eventually
run.
%
The last condition is required for TSNI; languages relying on
schedulers, such as |seqf|, that do not satisfy the low-progress
condition can at most satisfy TINI.
%
For simplicity, we do not treat this, more-general class of schedulers;
doing so would require modifying our calculus to keep track of the
label on the current label of tasks and complicate the
non-interference proofs.

More fundamentally, our scheduler parametrization is limited in being
deterministic, and thus extending the system to consider non-deterministic
schedulers (or even non-deterministic target languages) is non-trivial.
%
%For instance,  we do not consider  schedulers that, among other, rely
%on the global state |iS| or previous task configurations, as treated
%in~\cite{russo2006securing}.
%
%Additionally, our definitions rely on deterministic relations,
%
Such extensions often apply security conditions which suffer 
from refinement attacks (e.g.~\cite{Smith:Volpano:MultiThreaded}) or may require to
adapt non-trivial definitions as
\emph{low-determinism}~\cite{Zdancewic:Myers:CSFW03,Huisman:Worah:Sunesen:CSFW06,Terauchi:2008},
which states that a program is secure only if |l|-equivalent results are
deterministic.

@@

\begin{lemma}
  \label{lemma:high-not-blocking}
  For any task |it|, task lists |its|, store |iS|, and label |il|, if
  $|erase il it|=|bullet|$, then there exists a task list
  |its'| and a store |iS'| such that
  \begin{align}
  |iconf iS (it, its)| |.->| |iconf iS' (its, its')| \label{eq:hnb-1} \\
  |erase il its'|=|nil| \label{eq:hnb-2}\\
  |erase il iS'|=|erase il iS| \label{eq:hnb-3}
  \end{align}
\end{lemma}
\begin{proof}
  From $|erase il it|=|bullet|$ we know that the current label |lcurr| of |it|
  must be above |il|.  Furthermore, tasks can always take a step (if no
  regular rule applies, then \textsc{I-noStep} can be used), and thus
  we consider all rules that could be applied to execute |it|.
  \begin{description}
    \item[Case \textsc{I-noStep} and \textsc{I-done}]
    In this case, the task |it| is dropped,
    and thus |its'=nil| and |iS'=iS| satisfy
    conditions~\eqref{eq:hnb-2} and~\eqref{eq:hnb-3}.
    \item[Case \textsc{I-sandbox}]
    The newly created task has a label of at least |lcurr|, and will thus be
    erased, as required by condition~\eqref{eq:hnb-2}.  Furthermore, the
    state only changes for the newly created thread, and thus the state
    change is erased, showing~\eqref{eq:hnb-3}.
  \end{description}
  In all other rules, no new tasks are created, and thus |its'| consists of just
  the one task |it'|, to which |it| executed.  Since the tasks label can
  only increase, |it'| is still erased, showing condition~\eqref{eq:hnb-2}.
  We are left to show condition~\eqref{eq:hnb-3} for the remaining rules.
  \begin{description}
    \item[Case \textsc{I-send}]
    A new message triple with label |il'| gets added to the message
    queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
    the triple will get erased.
    \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
    In this case, only the queue of
    task |it| can change, which gets erased.
    \item[Case \textsc{I-new}] The newly allocated address has to be at a
    label at least as high as |lcurr|, and will thus be erased.
    \item[Case \textsc{I-write}] Only addresses with a label |il'| above
    |lcurr| can be written, thus the change in |iS1| will get erased.
    \item[Otherwise.]  None of the other rules modify the state |iS|, and
    thus |iS'=iS| will trivially satisfy condition~\eqref{eq:hnb-3}.
  \end{description}
\end{proof}

@@

>>>>

\section{Example IFC Language with a Single Heap}
\label{sec:single-heap}

@@

\section{Extending the Core Calculus}
\label{sec:appendix-extensions}

@@

%% \section{Limitations and Future Directions}
%% 
%% In this section we state some limitations of the current system, and
%% outline potential future directions to alleviate these limitations.
%% 
%% \subsection{Type Safety}
%% \label{sec:extensions:types}
%% 
%% One important consideration for any new language feature is whether
%% or not it preserves type safety.  If implementations are allowed to
%% have undefined behavior when the system gets stuck, concerns of type
%% safety can be directly applicable to security.
%% 
%% In our presentation, we have demanded that implementations \emph{not}
%% have undefined behavior when getting stuck: in such situations, the
%% \textsc{I-noStep} rule applies, where the stuck thread should be dropped
%% from execution.  Our combined language will be not get stuck, even if
%% the target language could get stuck.  This sort of guarantee can be
%% achieved at a coarse granularity for languages like C by using hard
%% isolation.  However, for many other languages, this is too stringent a
%% requirement.
%% 
%% A possible way to relax this requirement is to drop the \textsc{I-noStep}
%% rule, and instead demand type-safety of the combined language.  This
%% depends on the type system of the target language, so it is difficult to
%% make any general statements about how one would go about doing this.  However,
%% there are few general remarks to be made:
%% 
%% \begin{itemize}
%%     \item In Matthews and Findler's original
%%         paper~\cite{Matthews:2007:OSM:1190216.1190220}, an emphasis was
%%         on using type boundaries to mediate between the type systems of
%%         the two languages, i.e. handling conversions of values from
%%         one language to the other.  These techniques are directly applicable here.
%% 
%%     \item In the case of mini-ES, type safety proceeds by a preservation theorem
%%         that refers to well-typed stores.  In our setting, there are now multiple
%%         stores, and expressions may move from being interpreted with one store
%%         to another (e.g. when an address is sent from one thread to
%%         another); a clear violation of type safety.  In this case, operations
%%         such as |send| have their typing rules require
%%         the expressions being sent be typeable in an arbitrary well-typed store,
%%         statically disallowing the sending of addresses, or any types which
%%         may contain addresses.  In practice, this may be further restricted to
%%         types which can be easily marshalled, e.g. strings.
%% 
%%     \item Similarly, the choice of $\kappa$ (from Section~\ref{sec:retrofit}) directly
%%         influences the difficulty of the type-safety proof for
%%         |sandbox| when the target language has a store.  When
%%         $\kappa$ is identity, most expressions can be allowed, since the store
%%         remains the same; if $\kappa$ drops the store, however, only closed expressions
%%         can be sandboxed.
%% \end{itemize}
%% 
%% \subsection{Scheduling Policies}
%% Our specification language is parametrized by a scheduling policy |alpha|, which
%% maps a task list to a potentially different task list, i.e.,
%% $\alpha_{{\tiny\mathrm{\text{step}}}}(\ifc{\Varid{t}},\ldots)$. 
%% %
%% As previously discussed, the precise definition of |alpha|
%% dictates whether the language is TSNI, TINI, or neither.
%% %
%% We described two popular policies |roundrobinf| and
%% |seqf| for which we showed TSNI and TINI, respectively.
%% %
%% However, our definitions allow for other scheduling policies based on the
%% current task list---in fact, our formalism can be easily 
%% extended to consider history~\cite{BartheRRS07}.
%% %
%% As an example of generality, our approach allows to implement the scheduler of~\cite{Kashyap:2011}
%% which always schedules less sensitive threads first; this can be
%% employed to address external timing covert channels where an attacker
%% can measure the delay between output events.\footnote{
%%   When considering a label lattice that is not totally ordered, e.g.,
%%   DLM~\cite{myers:dlm} or DCLabels~\cite{dclabels}, this scheduling policy is
%%   considerably more complex.
%% }
%% %
%% More precisely, our calculus generalizes to a class of deterministic
%% schedulers that are: 1) \emph{integrity-preserving}, i.e., they do not
%% modify (or create) tasks; 2) \emph{non-interferent}, i.e., 
%% %greatest lower bound of all the labels on the task labels,\footnote{
%% %  The label on a task's current label is the current label of the
%% %  parent at the time of creation, i.e., the initial current label of
%% %  the task.
%% %} 
%% scheduling decisions on public tasks do not depend on a sensitive task's
%% state~\cite{russo2006securing}; and, 3) make \emph{low-progress}, i.e., if there
%% is a task in the task list whose label is below (or incomparable to) the label
%% of another task, this task (or another one at the same level) must eventually
%% run.
%% %
%% The last condition is required for TSNI; languages relying on
%% schedulers, such as |seqf|, that do not satisfy the low-progress
%% condition can at most satisfy TINI.
%% %
%% For simplicity, we do not treat this, more-general class of schedulers;
%% doing so would require modifying our calculus to keep track of the
%% label on the current label of tasks and complicate the
%% non-interference proofs.
%% 
%% More fundamentally, our scheduler parametrization is limited in being
%% deterministic, and thus extending the system to consider non-deterministic
%% schedulers (or even non-deterministic target languages) is non-trivial.
%% %
%% %For instance,  we do not consider  schedulers that, among other, rely
%% %on the global state |iS| or previous task configurations, as treated
%% %in~\cite{russo2006securing}.
%% %
%% %Additionally, our definitions rely on deterministic relations,
%% %
%% Such extensions often apply security conditions which suffer 
%% from refinement attacks (e.g.~\cite{Smith:Volpano:MultiThreaded}) or may require to
%% adapt non-trivial definitions as
%% \emph{low-determinism}~\cite{Zdancewic:Myers:CSFW03,Huisman:Worah:Sunesen:CSFW06,Terauchi:2008},
%% which states that a program is secure only if |l|-equivalent results are
%% deterministic.

@@

\begin{lemma}
  \label{lemma:high-not-blocking}
  For any task |it|, task lists |its|, store |iS|, and label |il|, if
  $|erase il it|=|bullet|$, then there exists a task list
  |its'| and a store |iS'| such that
  \begin{align}
  |iconf iS (it, its)| |.->| |iconf iS' (its, its')| \label{eq:hnb-1} \\
  |erase il its'|=|nil| \label{eq:hnb-2}\\
  |erase il iS'|=|erase il iS| \label{eq:hnb-3}
  \end{align}
\end{lemma}
\begin{proof}
  From $|erase il it|=|bullet|$ we know that the current label |lcurr| of |it|
  must be above |il|.  Furthermore, tasks can always take a step (if no
  regular rule applies, then \textsc{I-noStep} can be used), and thus
  we consider all rules that could be applied to execute |it|.
  \begin{description}
    \item[Case \textsc{I-noStep} and \textsc{I-done}]
    In this case, the task |it| is dropped,
    and thus |its'=nil| and |iS'=iS| satisfy
    conditions~\eqref{eq:hnb-2} and~\eqref{eq:hnb-3}.
    \item[Case \textsc{I-sandbox}]
    The newly created task has a label of at least |lcurr|, and will thus be
    erased, as required by condition~\eqref{eq:hnb-2}.  Furthermore, the
    state only changes for the newly created thread, and thus the state
    change is erased, showing~\eqref{eq:hnb-3}.
  \end{description}
  In all other rules, no new tasks are created, and thus |its'| consists of just
  the one task |it'|, to which |it| executed.  Since the tasks label can
  only increase, |it'| is still erased, showing condition~\eqref{eq:hnb-2}.
  We are left to show condition~\eqref{eq:hnb-3} for the remaining rules.
  \begin{description}
    \item[Case \textsc{I-send}]
    A new message triple with label |il'| gets added to the message
    queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
    the triple will get erased.
    \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
    In this case, only the queue of
    task |it| can change, which gets erased.
    \item[Case \textsc{I-new}] The newly allocated address has to be at a
    label at least as high as |lcurr|, and will thus be erased.
    \item[Case \textsc{I-write}] Only addresses with a label |il'| above
    |lcurr| can be written, thus the change in |iS1| will get erased.
    \item[Otherwise.]  None of the other rules modify the state |iS|, and
    thus |iS'=iS| will trivially satisfy condition~\eqref{eq:hnb-3}.
  \end{description}
  \qed
\end{proof}

@@

  \qed

### 2014-10-18 paper/conclusion.tex

In this paper, we argued that when designing a coarse-grained IFC
system, it is better to start with a fully isolated, multi-task system
and work one's way back to the model of a IFC language.  We showed
how to prove systems designed this way can be proved non-interferent,
without needing to rely on details of the target language, and
we provided conditions on how to securely refine our formal semantics to
consider optimizations required in practice.  We connected our semantics
to two IFC implementations for JavaScript based on this formalism,
%explained how our methodology improved an exiting IFC system for Haskell,
and proposed
an IFC system for C using hardware isolation.
By systematically applying ideas from IFC in operating systems to programming languages,
we hope to have elucidated some of the core design principles of coarse-grained,
dynamic IFC systems.

>>>>

In this paper, we argued that when designing a coarse-grained IFC
system, it is better to start with a fully isolated, multi-task system
and work one's way back to the model of a single language equipped
with IFC.  We showed
how to prove systems designed this way can be proved non-interferent
without needing to rely on details of the target language, and
we provided conditions on how to securely refine our formal semantics to
consider optimizations required in practice.  We connected our semantics
to two IFC implementations for JavaScript based on this formalism,
%explained how our methodology improved an exiting IFC system for Haskell,
and proposed
an IFC system for C using hardware isolation.
By systematically applying ideas from IFC in operating systems to programming languages,
we hope to have elucidated some of the core design principles of coarse-grained,
dynamic IFC systems.
